Rhyme: A Query Language for Nested Data Structures

### 2024-12-17 paper.tex
update figures and evaluation section

      \includegraphics[scale=0.5]{figures/evaluation.pdf}

@@

We also compared the execution time of \rhyme and TACO by running the generated code on four randomly generated sparse vectors with 10$\%$ sparsity. The results are shown in Figure \ref{evaluationpic}. As observed, \rhyme's generated code achieves approximately $86\%$ of TACO's performance for the example query.\par
The performance gap may stem from the fact that TACO generates native C code for the computation kernel, while \rhyme's backend relies on C++ and extensively uses STL containers. Additionally, \rhyme introduces several abstractions over both the sparse tensor data structure and the co-iteration method, which could contribute to additional overhead.\par
However, it is important to note that \rhyme is still an ongoing project. With minimal development effort, we could implement a pure C backend to generate more specialized C code. Furthermore, compared to TACO, \rhyme integrates relational algebra and sparse tensor algebra into a unified framework while supporting advanced loop optimizations such as loop hoisting and fusion.

>>>>

      \includegraphics[scale=0.7]{figures/evaluation.pdf}

@@

We also compared the execution time of \rhyme and TACO by running the generated code on four randomly generated sparse vectors with the same sparsity. The experiment was conducted with varying degrees of sparsity (10$\%$, 25$\%$, 50$\%$, and 75$\%$). Sparsity is defined as the ratio of non-zero values to all values. The results are shown in Figure \ref{evaluationpic}. As observed, \rhyme's generated code achieves approximately $90\%$ of TACO's performance at 10$\%$ sparsity for the example query. However, as sparsity increases, \rhyme's performance declines relative to TACO. At 75$\%$ sparsity, \rhyme achieves only $15\%$ of TACO's performance.\par
The performance gap may stem from several factors. First, TACO generates native C code for its computation kernel, whereas \rhyme's backend relies on C++ and heavily utilizes STL containers. Additionally, \rhyme introduces several abstractions over both the sparse tensor data structure and the co-iteration method implemented in the header file, which likely contribute to additional overhead. Moreover, \rhyme employs a different co-advancing algorithm than TACO. While TACO increments only the smallest pointer at each iteration, \rhyme identifies the largest pointer and advances all other pointers until their indices become greater than or equal to the current largest pointer. This may explain why the performance gap decreases as the vectors become sparser.\par
It is important to note that \rhyme is still an ongoing project, and performance is not the primary focus at this stage. With minimal development effort, we could implement a pure C backend to generate more specialized C code, reducing the overhead of C++ abstractions and STL containers. Additionally, we could adopt TACO's co-iteration logic as an alternative implementation in our header. Our primary contribution lies not in the already well-established area of sparse tensor computation kernels, but in efficiently integrating sparse tensor algebra within a declarative multi-paradigm language framework. Existing optimized co-iteration logic can easily be incorporated into our implementation. Furthermore, compared to TACO, \rhyme integrates relational algebra and sparse tensor algebra into a unified framework while supporting advanced loop optimizations, such as loop hoisting and fusion.
